### 2018-02-10 manuscript.Rnw
ypos et al

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-2.pdf}
	\caption{{\bf \texttt{behavr} table.}
	A: Illustration of a \texttt{behavr} object, core data structure in \texttt{rethomics}.
		The metadata holds a single row for each of the $n$ individuals. 
		Its columns, the $p$ metavariables, are of two kind: either required and defined by the acquisition platform (\emph{i.e.} used to fetch the data) or user-defined (\emph{i.e.} arbitrary).
		In the data, each row is a ``read'' (\emph{i.e.} information about one individual at one time point).
		It is formed of $q$ variables and is expected to have a very large number of reads $k$ for each individual $i$.
		Data and metadata are implicitly joined with the \texttt{id} field.
		Names used for variables and metavariable are only plausible examples that will likely differ in practice. 
		B: Non exhaustive list of uses of a \texttt{behavr} table, refered as \texttt{dt}. 
		In addition to operations on data, which are inherited from \texttt{data.table},
		we provide utilities designed specifically to act on both metadata and data.  
	}
	\label{fig:fig-2}
\end{figure}

@@

\section*{Availability and Future Directions}
All packages in the \texttt{rethomics} framework are available under the terms of the GPLv3 license and listed at
\href{https://github.com/rethomics}{https://github.com/rethomics/}.
Extensive installation instructions as well as reproducible demos and tutorials are available at
\href{https://rethomics.github.io/}{https://rethomics.github.io/}.
All packages are continuously integrated and unit tested on several version of \texttt{R} to minimise resent and future issues.

@@

Hannah

Patrick Kr{\"a}tschmer

Rob

Markus

>>>>

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-2.pdf}
	\caption{{\bf \texttt{behavr} table.}
	A: Illustration of a \texttt{behavr} object, the core data structure in \texttt{rethomics}.
		The metadata holds a single row for each of the $n$ individuals. 
		Its columns, the $p$ metavariables, are one of two kinds: either required -- and defined by the acquisition platform (\emph{i.e.} used to fetch the data) -- or user-defined (\emph{i.e.} arbitrary).
		In the data, each row is a ``read'' (\emph{i.e.} information about one individual at one time point).
		It is formed of $q$ variables and is expected to have a very large number of reads $k$ for each individual $i$.
		Data and metadata are implicitly joined with the \texttt{id} field.
		Note that the names used in this for variables and metavariable in this example are only plausible cases that will likely differ in practice. 
		B: Non exhaustive list of uses of a \texttt{behavr} table (refered as \texttt{dt}). 
		In addition to operations on data, which are inherited from \texttt{data.table},
		we provide utilities designed specifically to act on both metadata and data.  
	}
	\label{fig:fig-2}
\end{figure}

@@

\section*{Availability and Future Directions}
All packages in the \texttt{rethomics} framework are available under the terms of the GPLv3 license and listed at
\href{https://github.com/rethomics}{https://github.com/rethomics/}.
Extensive installation instructions as well as reproducible demos and tutorials are available at
\href{https://rethomics.github.io/}{https://rethomics.github.io/}.
All packages are continuously integrated and unit tested on several version of \texttt{R} to minimise the risk of present and future issues.

@@

Hannah, Alice, Diana

Markus, Patrick Kr{\"a}tschmer Rob

### 2018-02-10 manuscript.tex
ypos et al

\subsection*{Double plotted actograms}
``Double-plotted actograms'' are a common visualisation of periodicity and rhythmicity in circadian experiments.
In  \nameref{S1-Fig}A, we show the double-plotted actograms of each animals:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# we also show a subset of this figure in supplementary materials}
\hlstd{figS1A} \hlkwb{<-} \hlkwd{ggetho}\hlstd{(dt,} \hlkwd{aes}\hlstd{(}\hlkwc{z} \hlstd{= moving),} \hlkwc{multiplot} \hlstd{=} \hlnum{2}\hlstd{)} \hlopt{+}
            \hlcom{# bars n the z axis could }
            \hlcom{#  one could also use stat_tile_etho}
            \hlkwd{stat_bar_tile_etho}\hlstd{()} \hlopt{+}
            \hlcom{# divide plot by individual}
            \hlkwd{facet_wrap}\hlstd{(} \hlopt{~} \hlstd{label,} \hlkwc{ncol} \hlstd{=} \hlnum{4}\hlstd{)} \hlopt{+}
            \hlcom{# rename the y axis}
            \hlkwd{scale_y_discrete}\hlstd{(}\hlkwc{name}\hlstd{=}\hlstr{"Day"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\subsection*{Periodograms}
Ultimately, in order to quantify periodicity and rhythmicity, we compute periodograms.
In several methods are implemented in \texttt{zeitbebr}. In this example, we generate $\chi{}^2$ periodrograms and lay them out in a grid.
A representative sample of all animals is shown in Fig~\ref{fig:fig-4}B (see \nameref{S1-Fig}B for all individuals).

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{fig5A} \hlkwb{<-} \hlkwd{ggperio}\hlstd{(per_dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= power} \hlopt{-} \hlstd{signif_threshold,}
                             \hlkwc{colour} \hlstd{= genotype))} \hlopt{+}
          \hlcom{# periododogram a line for population mean}
          \hlcom{# and bootstrap error bars}
          \hlkwd{stat_pop_etho}\hlstd{(}\hlkwc{method} \hlstd{= ggplot2}\hlopt{::}\hlstd{mean_cl_boot)} \hlopt{+}
          \hlcom{# rename x and y axis }
          \hlkwd{scale_y_continuous}\hlstd{(}\hlkwc{name}\hlstd{=}\hlstr{"Relative power"}\hlstd{)} \hlopt{+}
          \hlkwd{scale_x_hours}\hlstd{(}\hlstr{"Period"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The lines shows and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak.
			      Individual animals are shown by dots whose size represent power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			C: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
			      }
	\label{fig:fig-5}
\end{figure}

@@

\texttt{summary\_dt} is just a regular data frame with one row per individual, containing both metadata and our summary statistics. It can therefore be used directly by \texttt{ggplot} and such:

@@

\texttt{R} provides one of the richest statistical toolbox available, which allows users to go deeper in he analysis of the extracted variables.
One could, for instance, perform a $\chi{}^2$ test on the number of rhythmic vs arhythmic flies in both genotypes.
To address the same question, we fit a binomial generalised linear model, which shows a strong positive effect of genotype Tim/NKCCOX on the probability of being rhythmic ($p$-value $< 10^{-5}$):
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{fit} \hlkwb{<-} \hlkwd{glm}\hlstd{(is_rhythmic} \hlopt{~} \hlstd{genotype, summary_dt,} \hlkwc{family}\hlstd{=}\hlstr{"binomial"}\hlstd{)}

>>>>

\subsection*{Double plotted actograms}
``Double-plotted actograms'' are a common visualisation of periodicity and rhythmicity in circadian experiments.
In  \nameref{S1-Fig}A, we show the double-plotted actograms of each animals. A representative sample of eight individuals is shown in Fig~\ref{fig:fig-4}A.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# we also show a subset of this figure in 4A}
\hlstd{figS1A} \hlkwb{<-} \hlkwd{ggetho}\hlstd{(dt,} \hlkwd{aes}\hlstd{(}\hlkwc{z} \hlstd{= moving),} \hlkwc{multiplot} \hlstd{=} \hlnum{2}\hlstd{)} \hlopt{+}
            \hlcom{# bars n the z axis could }
            \hlcom{#  one could also use stat_tile_etho}
            \hlkwd{stat_bar_tile_etho}\hlstd{()} \hlopt{+}
            \hlcom{# divide plot by individual}
            \hlkwd{facet_wrap}\hlstd{(} \hlopt{~} \hlstd{label,} \hlkwc{ncol} \hlstd{=} \hlnum{4}\hlstd{)} \hlopt{+}
            \hlcom{# rename the y axis}
            \hlkwd{scale_y_discrete}\hlstd{(}\hlkwc{name}\hlstd{=}\hlstr{"Day"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\subsection*{Periodograms}
Ultimately, in order to quantify periodicity and rhythmicity, we compute periodograms.
Several methods are implemented in \texttt{zeitbebr}. In this example, we generate $\chi{}^2$ periodrograms and lay them out in a grid.
A subset of eight animals is shown in Fig~\ref{fig:fig-4}B (see \nameref{S1-Fig}B for all individuals).

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{fig5A} \hlkwb{<-} \hlkwd{ggperio}\hlstd{(per_dt,} \hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= power} \hlopt{-} \hlstd{signif_threshold,}
                             \hlkwc{colour} \hlstd{= genotype))} \hlopt{+}
          \hlcom{# periododogram shown as a line for population mean}
          \hlcom{# and bootstrap error bars}
          \hlkwd{stat_pop_etho}\hlstd{(}\hlkwc{method} \hlstd{= ggplot2}\hlopt{::}\hlstd{mean_cl_boot)} \hlopt{+}
          \hlcom{# rename x and y axis }
          \hlkwd{scale_y_continuous}\hlstd{(}\hlkwc{name}\hlstd{=}\hlstr{"Relative power"}\hlstd{)} \hlopt{+}
          \hlkwd{scale_x_hours}\hlstd{(}\hlstr{"Period"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The solid lines and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak.
			      Individual animals are shown by dots whose size represent power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			C: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
			      }
	\label{fig:fig-5}
\end{figure}

@@

\texttt{summary\_dt} is just a regular data frame with one row per individual, containing both metadata and our summary statistics. It can therefore be used directly by \texttt{ggplot} and other tools:

@@

\texttt{R} provides one of the richest statistical toolbox available, which allows users to go deeper in he analysis of the extracted variables.
One could, for instance, perform a $\chi{}^2$ test on the number of rhythmic \emph{vs} arhythmic flies in both genotypes.
To address the same question, we fit a binomial generalised linear model, which shows a strong positive effect of genotype Tim/NKCCOX on the probability of being rhythmic ($p$-value $< 10^{-5}$):
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{fit} \hlkwb{<-} \hlkwd{glm}\hlstd{(is_rhythmic} \hlopt{~} \hlstd{genotype, summary_dt,} \hlkwc{family}\hlstd{=}\hlstr{"binomial"}\hlstd{)}

### 2018-02-10 results.Rnw
ypos et al

\subsection*{Double plotted actograms}
``Double-plotted actograms'' are a common visualisation of periodicity and rhythmicity in circadian experiments.
In  \nameref{S1-Fig}A, we show the double-plotted actograms of each animals:

<<cache=TRUE, warning=FALSE, result="hide", message=FALSE>>=
# we also show a subset of this figure in supplementary materials
figS1A <- ggetho(dt, aes(z = moving), multiplot = 2) +
            # bars n the z axis could 
            #  one could also use stat_tile_etho
            stat_bar_tile_etho() + 
            # divide plot by individual
            facet_wrap( ~ label, ncol = 4) +
            # rename the y axis
            scale_y_discrete(name="Day")
@

@@

ylabs <- ggplot_build(fig4A)$layout$panel_ranges[[1]]$y.labels
fig4A <- fig4A + scale_y_discrete(name="Day",
                                  limits = ifelse(1:length(ylabs) %%2, ylabs, ""))
@

\subsection*{Periodograms}
Ultimately, in order to quantify periodicity and rhythmicity, we compute periodograms.
In several methods are implemented in \texttt{zeitbebr}. In this example, we generate $\chi{}^2$ periodrograms and lay them out in a grid.
A representative sample of all animals is shown in Fig~\ref{fig:fig-4}B (see \nameref{S1-Fig}B for all individuals).

@@

<<cache=TRUE, warning=FALSE,message=FALSE, result="hide">>=
fig5A <- ggperio(per_dt, aes(y = power - signif_threshold,
                             colour = genotype)) + 
          # periododogram a line for population mean
          # and bootstrap error bars
          stat_pop_etho(method = ggplot2::mean_cl_boot) +
          # rename x and y axis 
          scale_y_continuous(name="Relative power") + 
          scale_x_hours("Period")
@

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The lines shows and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak.
			      Individual animals are shown by dots whose size represent power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			C: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
			      }
	\label{fig:fig-5}
\end{figure}

@@

\texttt{R} provides one of the richest statistical toolbox available, which allows users to go deeper in he analysis of the extracted variables.
One could, for instance, perform a $\chi{}^2$ test on the number of rhythmic vs arhythmic flies in both genotypes.
To address the same question, we fit a binomial generalised linear model, which shows a strong positive effect of genotype Tim/NKCCOX on the probability of being rhythmic ($p$-value $< 10^{-5}$):
<<>>=
fit <- glm(is_rhythmic ~ genotype, summary_dt, family="binomial")

>>>>

\subsection*{Double plotted actograms}
``Double-plotted actograms'' are a common visualisation of periodicity and rhythmicity in circadian experiments.
In  \nameref{S1-Fig}A, we show the double-plotted actograms of each animals. A representative sample of eight individuals is shown in Fig~\ref{fig:fig-4}A.

<<cache=TRUE, warning=FALSE, result="hide", message=FALSE>>=
# we also show a subset of this figure in 4A
figS1A <- ggetho(dt, aes(z = moving), multiplot = 2) +
            # bars n the z axis could 
            #  one could also use stat_tile_etho
            stat_bar_tile_etho() + 
            # divide plot by individual
            facet_wrap( ~ label, ncol = 4) +
            # rename the y axis
            scale_y_discrete(name="Day")
@

@@

ylabs <- ggplot_build(fig4A)$layout$panel_ranges[[1]]$y.labels

new_labs <- ifelse(1:length(ylabs) %%2, ylabs, "")

fig4A <- fig4A + scale_y_discrete(name="Day",
                                  labels=new_labs
                                  )
@

\subsection*{Periodograms}
Ultimately, in order to quantify periodicity and rhythmicity, we compute periodograms.
Several methods are implemented in \texttt{zeitbebr}. In this example, we generate $\chi{}^2$ periodrograms and lay them out in a grid.
A subset of eight animals is shown in Fig~\ref{fig:fig-4}B (see \nameref{S1-Fig}B for all individuals).

@@

<<cache=TRUE, warning=FALSE,message=FALSE, result="hide">>=
fig5A <- ggperio(per_dt, aes(y = power - signif_threshold,
                             colour = genotype)) + 
          # periododogram shown as a line for population mean
          # and bootstrap error bars
          stat_pop_etho(method = ggplot2::mean_cl_boot) +
          # rename x and y axis 
          scale_y_continuous(name="Relative power") + 
          scale_x_hours("Period")
@

\begin{figure}[!h]
	%\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The solid lines and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak.
			      Individual animals are shown by dots whose size represent power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			C: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
			      }
	\label{fig:fig-5}
\end{figure}

@@

\texttt{R} provides one of the richest statistical toolbox available, which allows users to go deeper in he analysis of the extracted variables.
One could, for instance, perform a $\chi{}^2$ test on the number of rhythmic \emph{vs} arhythmic flies in both genotypes.
To address the same question, we fit a binomial generalised linear model, which shows a strong positive effect of genotype Tim/NKCCOX on the probability of being rhythmic ($p$-value $< 10^{-5}$):
<<>>=
fit <- glm(is_rhythmic ~ genotype, summary_dt, family="binomial")
